Low field depolarization of NV$^-$ centers through dipole-dipole coupling

### 2022-06-02 Mes papiers/Low field CR/Low_field_CR.tex
tata yoyo

\title{Low field depolarization of electronic spins through dipole-dipole coupling}

\author{C. Pellet-Mary$^1$, M. Perdriat$^1$, G. H\'etet}

@@

The ensemble spin lifetime is non-exponential due to inhomogeneities between the NV$^-$ centers. We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi_depolarization_2017}. A conclusion of this model is that, for an homogeneous 3D distribution of fluctuator, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$. Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$  (see SI).

For the samples used in this paper (detailed in SI), we observed that $T_1^{\rm dd} \sim T_1^{\rm ph}$, meaning that we had to include both  in our analysis. The resulting fitting formula is therefore :
\begin{equation}
S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}}),
\end{equation}
In order to get comparable fitting parameters between each measurement, we decided to fix the value of the exponential lifetime in the fitting process, leaving only the stretch lifetime as a free parameter. For Fig.\ref{T1}, \ref{largeur_fluct}, \ref{100_VS_1x4} and \ref{exp_B_transverse}, the chosen exponential lifetime was $T_1^{\rm ph}=3.62 \ \rm ms$.

\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
\caption{Tuning resonant coupling through magnetic field orientation. (a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice. (b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes. The detuning $\Delta \nu$ between the classes $\beta$ and $\gamma$ was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_\gamma$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz . Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
\label{largeur_fluct}
\end{figure}

Fig. \ref{largeur_fluct} illustrates how we can use the NV center geometry to tune the resonant dipolar-coupling between the spins. Because of the diamond symmetry, we can distinguish 4 groups ("classes") of NV centers, based on the four possible orientation of the NV axis in the crystal. These four classes have \textit{a priori} different spin resonance frequency based on the projection of the magnetic field on their respective NV axis. By tuning the angle of the magnetic field, we can therefore control the frequency detuning between the different classes, and as a result we can control the number of resonant NV centers by bringing in and out of resonance several classes.

Fig. \ref{largeur_fluct} (c) in particular shows the dipole-induced lifetime $T_1^{\rm dd}$ for a given class as a function of the energy detuning with another class. One can notice that the width of the $T_1^{\rm dd}$ bell shape is significantly larger than that of the frequency overlap between the two classes, modeled here by an ODMR line stretched by a factor $\sqrt{2}$ (see SI). This extra width can not be attributed to the dipole-dipole coupling strength : for a 3 ppm concentration of NV$^-$ centers, the average coupling strength between neighboring NV centers should be $\sim 27\ \rm kHz$, two order of magnitude below the observed width. We can also see that $T_1^{\rm dd}$ follows a Lorentzian shape despite the ODMR lines being non-Lorentzian. These two observations corroborate the idea developed in \cite{choi_depolarization_2017} that the interaction width is dominated by the noise spectrum of the lifetime-limited fluctuators.

@@

The depolarization in low field can be explained in large part due to the lift of degeneracy between the four classes, as can be seen in Fig. \ref{100_VS_1x4} : On the top part, a magnetic field is scanned in a direction chosen in order to quickly split the transition frequencies of the four NV classes. We can observe an increase of about 4\% in the photoluminescence and an increase by a factor of $\sim$ 16 in the spin dipolar-induced lifetime ${T_1^{\rm dd}}$ as the magnetic field is increased.

However the splitting of the four classes can not explain entirely the spin depolarization in zero field. The bottom part of Fig. \ref{100_VS_1x4} shows a similar experiment when the magnetic field is aligned with the [100] crystalline axis. In this particular orientation, the projection of the magnetic field on the four NV classes is identical, meaning that the resonance condition between the NV centers remains the same in zero and non-zero field. Yet we can still observe an increase of $\sim$ 0.5\% in the photoluminescence and an increase of $\sim$ 30\% in ${T_1^{\rm dd}}$.
We attribute the remaining depolarization to two phenomena : the change of basis in the spin Hamiltonian's eigenstate in zero-field, and the near-resonance of the dipolar double-flip processes.

The decrease of spin lifetime and the associated drop in PL for $B\approx 20\ \rm G$ is due to cross-relaxations between NV $^-$ centers with a $^{13}$C as first a neighbor, and those without \cite{pellet2021optical}.

The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :
\begin{equation}
\mathcal{H}^{\rm dd}= -\frac{J_0}{r^3}\left(3\left({\vec S}_1 \cdot \vec u \right)\left({\vec S}_2 \cdot \vec u \right) - {\vec S}_1 \cdot {\vec S}_2  \right),
\end{equation}
Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$. In situations such as the one described in Fig. \ref{largeur_fluct}, the only relevant terms of $\mathcal{H}^{\rm dd}$ in term of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$. In zero magnetic field however, we have to take into account other aspects of the dipolar Hamiltonian in consideration.

>>>>

\title{Low field depolarization of NV$^-$ centers through dipole-dipole coupling}

\author{C. Pellet-Mary$^1$, P. Huillery$^?$,  M. Perdriat$^1$, G. H\'etet}

@@

The ensemble spin decay is non-exponential due to inhomogeneities between the NV$^-$ centers. We will base our interpretations of the experimental results on the NV-fluctuator model developed in \citep{choi_depolarization_2017}. A conclusion of this model is that, for an homogeneous 3D distribution of fluctuator, the dipole-induced lifetime should be stretched-exponential with a stretch factor $\beta=1/2$. Such an ensemble lifetime is indeed observed when $T_1^{\rm dd} \ll T_1^{\rm ph}$  (see SI).

For the samples used in this paper (detailed in SI), we observed that $T_1^{\rm dd} \sim T_1^{\rm ph}$, meaning that we had to include both  in our analysis. The resulting fitting formula is therefore :
\begin{equation}
S(\tau)=A \exp (-\frac{\tau}{T_1^{\rm ph}} -\sqrt{\frac{\tau}{T_1^{\rm dd}}}),
\end{equation}
Further details on the fitting procedure are available in SI. For the rest of the article we will be only interested on the stretched exponential lifetime ${T_1^{\rm dd}}$.

%\begin{figure}
%\includegraphics[width=0.45\textwidth]{Figures/fig largeur fluct}
%\caption{Tuning resonant coupling through magnetic field orientation. (a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice. (b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes. The detuning $\Delta \nu$ between the classes $\beta$ and $\gamma$ was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_\gamma$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz . Green dashed line correspond to the ODMR width of a single class stretched by a factor $\sqrt 2$, approximating the overlap of the two classes (see SI).}
%\label{largeur_fluct}
%\end{figure}

%Fig. \ref{largeur_fluct} illustrates how we can use the NV center geometry to tune the resonant dipolar-coupling between the spins. Because of the diamond symmetry, we can distinguish 4 groups ("classes") of NV centers, based on the four possible orientation of the NV axis in the crystal. These four classes have \textit{a priori} different spin resonance frequency based on the projection of the magnetic field on their respective NV axis. By tuning the angle of the magnetic field, we can therefore control the frequency detuning between the different classes, and as a result we can control the number of resonant NV centers by bringing in and out of resonance several classes.
%
%Fig. \ref{largeur_fluct} (c) in particular shows the dipole-induced lifetime $T_1^{\rm dd}$ for a given class as a function of the energy detuning with another class. One can notice that the width of the $T_1^{\rm dd}$ bell shape is significantly larger than that of the frequency overlap between the two classes, modeled here by an ODMR line stretched by a factor $\sqrt{2}$ (see SI). This extra width can not be attributed to the dipole-dipole coupling strength : for a 3 ppm concentration of NV$^-$ centers, the average coupling strength between neighboring NV centers should be $\sim 27\ \rm kHz$, two order of magnitude below the observed width. We can also see that $T_1^{\rm dd}$ follows a Lorentzian shape despite the ODMR lines being non-Lorentzian. These two observations corroborate the idea developed in \cite{choi_depolarization_2017} that the interaction width is dominated by the noise spectrum of the lifetime-limited fluctuators.

@@

One of the way we can isolate dipole-dipole related phenomena with NV centers is by tuning the number of NV centers resonant with each other. Indeed, Fig. \ref{100_VS_1x4} (bi) shows the characteristic 8 lines observed with ensemble NV optically detected magnetic resonance (ODMR) spectrum. These 8 lines correspond to the four "classes" of NV centers - that is physical orientation of the NV axis is the diamond crystal cell - multiplied by the two possible spin transitions $\ket{0}\to\ket{-1}$ and $\ket{0}\to\ket{+1}$. The transitions of the four different classes can be controlled somewhat independently by changing the amplitude and orientation of the external magnetic field, and in particular they can be brought to resonance for particular orientation of the magnetic field. Fig. \ref{100_VS_1x4} (ci) shows an example where all four classes are brought to resonance by placing the magnetic field along the [100] crystalline axis. Other example of class resonance are given in SI.

Fig. \ref{100_VS_1x4} (bii) and (biii) shows respectively the PL and the stretched exponential spin decay $T_1^{\rm dd}$ of an ensemble of NV centers with respect to the amplitude of the external magnetic field. The orientation of the magnetic field was chosen in order to lift the degeneracy between the four classes, as shown in Fig. \ref{100_VS_1x4} (bi). We observe that the drop in PL at low magnetic field is indeed correlated with an increase in the spin decay rate, as is expected due to the higher number of resonant spins in zero magnetic field, where all four classes are resonant, compared to the non-degenerate case for $\vec \neq 0$. We can also see that the decrease in PL for $B>40\ \rm G$ due to state mixing by the transverse magnetic field is not correlated to a modification of the spin lifetime.

However, the splitting of the four classes is not sole reason to the decrease of the spin lifetime in low magnetic field. Indeed, we performed similar measurements in Fig. \ref{100_VS_1x4} (cii) and (ciii) but this time fixing the magnetic field orientation along the [100] axis. For this particular orientation, the four classes are always resonant, regardless of the field amplitude. We can see that, although considerably reduced, there still is an increase in the spin decay rate and a corresponding drop in PL for low magnetic field values. We attribute the slight drop in PL and the corresponding bump for $1/T_1^{\rm dd}$ at $B \sim 20\ \rm G$ to dipolar interaction with NV having a first neighbor $^{13}C$ \cite{pellet2021optical}.

The aim of this paper is to understand the remaining contribution to the spin decay in this second case. We identified and isolated two additional effects : the domination by the local electric field for low magnetic field value which results in a change of the spin Hamiltonian eigenstates, and the presence of double-flip processes where both NV spins involved in the dipole-dipole coupling flip their spin in the same direction.

In order to distinguish these two contributions, we decided to exploit the transverse magnetic field.... (lis un peu ce que t'as déja écris avant de tout réécrire, je ter connais moi futur)
The dipole-dipole interaction Hamiltonian between two spins ${\vec S}_1$ and ${\vec S}_2$ reads :
\begin{equation}
\mathcal{H}^{\rm dd}= -\frac{J_0}{r^3}\left(3\left({\vec S}_1 \cdot \vec u \right)\left({\vec S}_2 \cdot \vec u \right) - {\vec S}_1 \cdot {\vec S}_2  \right),
\end{equation}
Where $J_0= (2 \pi) 52\ \rm MHz \cdot \rm{nm}^3$, $\vec r$ is the relative positions of the two spins and $\vec u = \frac{\vec r}{\norm{\vec r}}$. In situations such as the one described in Fig. \ref{largeur_fluct}, the only relevant terms of $\mathcal{H}^{\rm dd}$ in term of population transfer are the flip-flop terms such as $\mel{0,+1}{\mathcal{H}^{\rm dd}}{+1,0}$. In zero magnetic field however, we have to take into account other aspects of the dipolar Hamiltonian in consideration.

### 2022-06-02 Mes papiers/Low field CR/SI/Low_field_CR_SI.tex
tata yoyo

\begin{document}
\vspace{0.2in}
{\Large \hspace{1.6in}\textsc{Supplementary Material} }\\
\
\title{Low field depolarization of electronic spins through dipole-dipole coupling}

@@

An other possibility would be to use an arbitrary stretched factor in the fitting function : $S(\tau)=A \exp (-(\frac{\tau}{T_1^{\rm ph}})^\beta )$. Fig. \ref{alphas} shows the optimal $\beta$ parameter as a function of the external magnetic field, which confirms that the $T_1$ profile gets closer to a pure stretched-exponential in zero magnetic field.
\begin{figure}
\includegraphics[width=0.8\textwidth]{Figures_SI/Fig_T1_combined}
\caption{$T_1$ measurement with purely exponential and purely stretched exponential fits (a) in zero magnetic field (b) in non-zero magnetic field}
\label{T1_fits}
\end{figure}
\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures_SI/fig_alphas}
\caption{Best stretch factor $\beta$ for a $T_1$ fit of the form $f(\tau)=A \exp(-(\frac{\tau}{T_1})^\beta)$ as a function of a randomly oriented magnetic field amplitude}
\label{alphas}
\end{figure}
\subsection{Spectral range of the dipole-dipole cross-relaxations}
\begin{figure}
\includegraphics[width=0.7\textwidth]{Figures_SI/largeur_fluct_SI}
\caption{Dipole-dipole depolarization for two near-resonant classes. (a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice. (b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes. The detuning $\Delta \nu$ between the classes b and c was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_c$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz. Single class ESR line stretched by a factor of 1,$\sqrt{2}$ and 2 are added for comparison}
\label{largeur_fluct}
\end{figure}
In the absence of other

@@

These predictions can be tested experimentally. Fig. \ref{121 VS 22 fig}(b) shows the subtracted $T_1$ signal measured on the same sample [REF] for 8 different values of the magnetic field, each time on a two-class resonance : in 4 cases the magnetic field was in a \{100\} plane, and in the 4 other cases it was in a \{110\} plane. We can see that the measured lifetime is always smaller on the \{110\} case, which corresponds the the greater $\bar \eta$ factor computed previously.

@@

Here is a list of the computed values of $\bar \eta^2$, which according to eq.(\ref{eq 1/T}) is proportional to the dipole induced decay rate, for various magnetic field orientation. An ODMR spectrum for each of these situations is present in Fig. \ref{Various ODMR}.
\begin{itemize}
\item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$
\item $\vec{B} \in$ \{110\} : For the two non resonant classes : $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$.  For the two resonant classes : $\bar \eta = \bar {\eta}_{same} + \bar {\eta}_{far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$
\item $\vec{B} \in$ \{100\} : For each pair of two-classes resonance: $\bar \eta = \bar {\eta}_{same}+\bar {\eta}_{close}$ and $\bar \eta^2\approx 7.24\, \bar \eta_0^2$. 
\item $\vec{B} \parallel \langle 111 \rangle$ : For the non-resonant class (parallel to $\vec B$) : $\bar \eta = \bar {\eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$. For the triply resonant classes: $\bar \eta = \bar {\eta}_{same}+2\bar {\eta}_{far}$ and $\bar \eta^2\approx 28.4\, \bar \eta_0^2$. 
\item $\vec{B} \parallel \langle 100 \rangle$ : For the quadruple resonance : $\bar \eta = \bar {\eta}_{same}+2\bar {\eta}_{far}$ and $\bar \eta^2\approx 42.8\, \bar \eta_0^2$.
\end{itemize}

The measured increase in the decay rate is generally smaller than the predicted one : we measured an increase by a factor $\sim 4$ for a two class degeneracy instead of the predicted $10.0$, and an increase by a factor $\sim 16$ for a two class degeneracy instead of the predicted $42.8$. \cite{choi_depolarization_2017} measured an increase by a factor $\sim 4$ on a two-class degeneracy instead of the predicted $7.24$ or $10.0$.

@@

As described in sec. \ref{sec Hamiltonian}, when the single spin Hamiltonian of the NV center is dominated by the electric field or the transverse magnetic field (as long as $\gamma_e B_\perp \ll D$), the eigen basis of the NV Hamiltonian is close to $\{\ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $. We therefore have to rewrite the dipole-dipole Hamiltonian using this new basis for both spins in order to perform similar calculations as the ones done in the previous section.

@@

\bibliographystyle{plain}
\bibliography{CR_SI}{}
\end{document}

>>>>

\begin{document}
\vspace{0.2in}
{\Large \hspace{1.6in}\textsc{Supplementary Material} }\\
\
\title{Low field depolarization of NV$^-$ centers through dipole-dipole coupling}

@@

An other possibility would be to use an arbitrary stretched factor in the fitting function : $S(\tau)=A \exp (-(\frac{\tau}{T_1^{\rm ph}})^\beta )$. Fig. \ref{alphas} shows the optimal $\beta$ parameter as a function of the external magnetic field, which confirms that the $T_1$ profile gets closer to a pure stretched-exponential in zero magnetic field.
\begin{figure}
\includegraphics[width=0.8\textwidth]{Figures_SI/Fig_T1_combined}
\caption{$T_1$ measurement with purely exponential and purely stretched exponential fits (a) in zero magnetic field (b) in non-zero magnetic field}
\label{T1_fits}
\end{figure}
\begin{figure}
\includegraphics[width=0.45\textwidth]{Figures_SI/fig_alphas}
\caption{Best stretch factor $\beta$ for a $T_1$ fit of the form $f(\tau)=A \exp(-(\frac{\tau}{T_1})^\beta)$ as a function of a randomly oriented magnetic field amplitude}
\label{alphas}
\end{figure}
\subsection{Spectral range of the dipole-dipole cross-relaxations}
\label{fluctuator width}
\begin{figure}
\includegraphics[width=0.7\textwidth]{Figures_SI/largeur_fluct_SI}
\caption{Dipole-dipole depolarization for two near-resonant classes. (a) Sketch of the four possible orientations ("classes") in a single crystal diamond lattice. (b) ODMR spectrum showing four $\ket{0} \to \ket{-1}$ resonances corresponding to the four spin classes. The detuning $\Delta \nu$ between the classes b and c was controlled by changing the orientation of the external magnetic field.(c) Stretch part of the lifetime decay for the spins resonant with $\nu_c$ as a function of the detuning $\Delta \nu$ (blue circles), fitted by a Lorentzian with half width at half maximum 8.04 MHz. Single class ESR line stretched by a factor of 1,$\sqrt{2}$ and 2 are added for comparison}
\label{largeur_fluct}
\end{figure}

@@

These computed values can be tested experimentally. Fig. \ref{121 VS 22 fig}(b) shows the subtracted $T_1$ signal measured on the same sample [REF] for 8 different values of the magnetic field, each time on a two-class resonance : in 4 cases the magnetic field was in a \{100\} plane, and in the 4 other cases it was in a \{110\} plane. We can see that the measured lifetime is always smaller on the \{110\} case, which corresponds the the greater $\bar \eta$ factor computed previously.

@@

Here is a list of the computed values of $\bar \eta^2$, which according to eq.(\ref{eq 1/T}) is proportional to the dipole induced decay rate, for various magnetic field orientation. An ODMR spectrum for each of these situations is present in Fig. \ref{Various ODMR}.
\begin{itemize}
\item Random orientation / no class degeneracy : for all four classes, $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2=3.08 \cdot 10^{-3} \equiv \bar \eta_0^2$
\item $\vec{B} \in$ \{110\} : For the two non resonant classes : $\bar \eta = \bar {\eta}_{\rm same}$ and $\bar \eta^2= \bar \eta_0^2$.  For the two resonant classes : $\bar \eta = \bar {\eta}_{\rm same} + \bar {\eta}_{\rm far}$ and $\bar \eta^2\approx\, 10.0 \bar \eta_0^2$
\item $\vec{B} \in$ \{100\} : For each pair of two-classes resonance: $\bar \eta = \bar {\eta}_{\rm same}+\bar {\eta}_{\rm close}$ and $\bar \eta^2\approx 7.24\, \bar \eta_0^2$. 
\item $\vec{B} \parallel \langle 111 \rangle$ : For the non-resonant class (parallel to $\vec B$) : $\bar \eta = \bar {\rm \eta}_{same}$ and $\bar \eta^2= \bar \eta_0^2$. For the triply resonant classes: $\bar \eta = \bar {\eta}_{\rm same}+2\bar {\eta}_{\rm far}$ and $\bar \eta^2\approx 28.4\, \bar \eta_0^2$. 
\item $\vec{B} \parallel \langle 100 \rangle$ : For the quadruple resonance : $\bar \eta = \bar {\eta}_{\rm same}+2\bar {\eta}_{\rm close} + \bar {\eta}_{\rm far}$  and $\bar \eta^2\approx 42.8\, \bar \eta_0^2$.
\end{itemize}

The measured increase in the decay rate is generally smaller than the predicted one : we measured an increase by a factor $\sim 4$ for a two-class degeneracy instead of the predicted $10.0$, and an increase by a factor $\sim 16$ for a four-class degeneracy instead of the predicted $42.8$. \cite{choi_depolarization_2017} measured an increase by a factor $\sim 4$ on a two-class degeneracy instead of the predicted $7.24$ or $10.0$.

@@

As described in sec. \ref{sec Hamiltonian}, when the single spin Hamiltonian of the NV center is dominated by the electric field or the transverse magnetic field (as long as $\gamma_e B_\perp \ll D$), the eigenbasis of the NV Hamiltonian is close to $\{\ket{0},\ket{+}=\frac{\ket{+1}+\ket{-1}}{\sqrt{2}},\ket{-}=\frac{\ket{+1}-\ket{-1}}{\sqrt{2}} \} $. We therefore have to rewrite the dipole-dipole Hamiltonian using this new basis for both spins in order to perform similar calculations as the ones done in the previous section.

Since the flip-flop processes are now of the form $\ket{0,\pm}\bra{\pm,0}$, we redefine the $\eta$ factor as :
\begin{equation}
\eta^2=\frac{1}{3} \abs{\mel{\pm ,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,\pm }}^2  \frac{4\gamma_f^2}{(\omega_f - \omega_{NV})^2+4\gamma_f^2}
\end{equation}

@@

\begin{equation}
  S_x = \begin{pmatrix}
  0&1&0 \\
  1&0&0 \\
  0&0&0
  \end{pmatrix}
  \begin{matrix}
  \bra{-} \\
  \bra{0} \\
  \bra{+}
  \end{matrix}  
  , \quad 
 S_y = \begin{pmatrix}
  0&0&0 \\
  0&0&1 \\
  0&1&0
  \end{pmatrix}
  \begin{matrix}
  \bra{-} \\
  \bra{0} \\
  \bra{+}
  \end{matrix}  
 , \quad 
  S_z = \begin{pmatrix}
  0&0&1 \\
  0&0&0\\
  1&0&0
  \end{pmatrix}
  \begin{matrix}
  \bra{-} \\
  \bra{0} \\
  \bra{+}
  \end{matrix}
  \end{equation}

The symmetry in the $(xy)$ plane is broken by the presence of the transverse electric or magnetic field. The $x_i$ direction is defined as the direction of the transverse electric or magnetic field which dominates the single spin Hamiltonian of the particle $i$.

We will consider two cases :
\begin{itemize}
\item The $x$ direction is defined by an external field (transverse magnetic field or strong external electric field) projected on the $(xy)$ plane of each classes. In particular this means that two spins from the same class have the same $(\hat x, \hat y, \hat z)$ basis.
\item The $x$ direction is defined by the local electric field generated by the charges in the crystal. We will assume that there are no correlations in the electric field between the NV and fluctuator, and will sample a random angle $\psi \in [0,2\pi]$ between the axes $\hat{x}_1$ and $\hat{x}_2$.
\end{itemize}

\begin{table}
\begin{tabular}{cccc}
\hline
$\bar{\eta}$ table & $\widehat{z_1 z_2}=0$ & $\widehat{z_1 z_2}=70.5^\circ$ & $\widehat{z_1 z_2}=109.5^\circ$ \\
\hline
$\ket{\pm 1}$ basis & $\frac{2}{3\sqrt{3}}=$ 0.3849 & 0.6507 & 0.8328 \\
$\ket{+/-}$ basis $\hat{x}_1\neq \hat{x}_2$ & 0.7110  & 0.6828 & 0.6828 \\
$\ket{+/-}$ basis $\hat{x}_1= \hat{x}_2$ & $\frac{4}{3\sqrt{3}}=$0.7698  & 0.6951 & 0.6951 \\
\hline
\end{tabular}
\caption{Computation of $\frac{\bar{\eta}}{ \frac{1}{4} \cdot \sqrt{\frac{1}{3}}}$ for the different eignebasis of the single spin Hamiltonian, and for the different angles between the $z$ axis of the two spins.}
\label{table eta}
\end{table}

For this two cases, which we will call $\hat{x}_1=\hat{x}_2$ and $\hat{x}_1\neq \hat{x}_2$, we can compute the $\bar \eta$ factor for the three scenarios discussed previously : $\widehat{z_1 z_2}=0$ (same class), $\widehat{z_1 z_2}=70.5^\circ$ and $\widehat{z_1 z_2}=109.5^\circ$. The results, as well as those in the $\ket{\pm 1}$ basis are presented in Table \ref{table eta}.

There are two particular situations where these values can be experimentally tested :

The first one is described in Fig. [REF] of the main text : in the presence of pure transverse magnetic field (for a single class, non-resonant with the three other classes), the decay rate of the class decreases with the magnetic field amplitude, due to the double-flip processes (described below), and then reaches a plateau with a decay rate value $\sim 2$ times bigger than in the longitudinal magnetic field case. This increase is in agreement with the fact that $\bar{\eta}_{\rm same}$ is 2 times bigger in the non-magnetic $\ket{+/-}$ basis than in the magnetic $\ket{\pm 1}$ basis. The measured increase is again smaller than the predicted factor of 4 given by eq. (\ref{eq 1/T}).

The second one is the difference between the $\vec{B} \parallel \langle 100 \rangle$ case and the $\vec{B}=0$ case. When a magnetic field is applied in the [100] direction, all four classes are resonant and the spin Hamiltonian basis is $\{ \ket{0},\ket{+1},\ket{-1} \} $. The $\bar \eta$ factor in this case has been calculated in the previous section : $\bar \eta^2\approx 42.8\, \bar \eta_0^2$. When no external magnetic field is applied, all four classes are also resonant but the spin Hamiltonian basis is  $\{ \ket{0},\ket{+},\ket{-} \} $ (see Sec. \ref{NV Hamiltonian}) and double-flip processes are near-resonant. The decay rate due purely to flip-flop in the $\ket{+/-}$ basis is proportional to $\bar \eta^2 = (\bar{\eta}_{\rm same}^{+/-} + 3\bar{\eta}_{\rm diff}^{+/-})^2 \approx 51.4\, \bar \eta_0^2$. Fig. [REF] of the main text shows that we indeed observe a shorter lifetime when $\vec B=0$, in agreement with the higher value of $\bar \eta^2$, but this decrease could also come from the double-flip processes which are absent when $\vec B \neq 0$.

\subsection{Double-flip processes}

Double-flip processes, that is processes where the two spins flip in the same direction, compared to flip-flop where the two spins flip in opposite directions, are related to matrix elements  such as $\mel{+1,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-1}$ in the $\ket{\pm 1}$ basis or $\mel{+,0}{\frac{\mathcal{H}_{\rm dd}}{J_0/r^3}}{0,-}$ in the $\ket{\pm}$ basis. These matrix elements couple two-spin states that are never fully resonant, however with low-to zero magnetic, the residual splitting due to local electric and magnetic field ($\sim 6\ \rm MHz$ with our samples) is small enough compared to the fluctuators linewidth  measured in sec. \ref{fluctuator width} to be $\sim 8\ \rm MHz$, so that double-flip processes may still occur.

Fig. [REF] of the main text shows that for lower transverse field values, the spin decay rate is increased significantly. We attribute this increase to double-flip processes. We can see that the decay rate increase ($\sim$ 5 times the baseline value) is significantly higher than the increase due solely to the $\ket{+/-}$ basis ($\sim$ 2 times the baseline value). This lead us to believe that the double-flip process are the main reason behind the decrease in spin lifetime for $\vec B=0$ observed in Fig. [REF] of the main text. This would probably mean that double-flip processes are the dominant factor in the sensitivity of the magnetometry protocol presented in the main text.
\bibliographystyle{plain}
\bibliography{CR_SI}{}
\end{document}
